---
title: "Lab3"
author: "Jayson De La O"
format:
  html:
    embed-resources: true
---

## Quarto
```{r}
library(dplyr)
library(leaflet)
library(devtools)
```

Code to download dataset
download.file(
  "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
  destfile = file.path("~", "Downloads", "met_all.gz"),
  method   = "libcurl",
  timeout  = 60
)
```{r}
met <- data.table::fread(file.path("~", "Downloads", "met_all.gz"))
```

```{r}
dim(met)
head(met)
tail(met)
```

```{r}
str(met)
table(met$year)
table(met$day)
table(met$hour)
```

```{r}
summary(met$temp)
summary(met$elev)
summary(met$wind.sp)
```

```{r}
met[met$elev==9999.0] <- NA
summary(met$elev)
```

```{r}
met <- met[temp>-40]
met2 <- met[order(temp)]
head(met2)
```
Temps higher than -40

```{r}
summary(met$temp)
```
temps higher than -17
```{r}
met <- met[temp>-15]
met2 <- met[order(temp)]
head(met2)
summary(met$temp)
```



```{r}
elev <- met[elev==max(elev), ]
summary(elev)
```


```{r}
cor(elev$temp, elev$wind.sp, use="complete")
cor(elev$temp, elev$hour, use="complete")
cor(elev$wind.sp, elev$day, use="complete")
cor(elev$wind.sp, elev$hour, use="complete")
cor(elev$temp, elev$day, use="complete")
```



```{r}
hist(met$elev, breaks=100)
hist(met$temp)
hist(met$wind.sp)
```



```{r}
leaflet(elev) %>%
  addProviderTiles('OpenStreetMap') %>% 
  addCircles(lat=~lat,lng=~lon, opacity=1, fillOpacity=1, radius=100)
```

```{r}
library(lubridate)
elev$date <- with(elev, ymd_h(paste(year, month, day, hour, sep= ' ')))
summary(elev$date)
```

```{r}
elev <- elev[order(date)]
head(elev)
```

```{r}
plot(elev$date, elev$temp, type='l')
```

```{r}
plot(elev$date, elev$wind.sp, type='l')
```

Trends in the time series plots: There are rapid increases and decreases in both plots which seems reasonable because temperatures rise throughout the day and cool during the night. The winds are always increasing which we would expect for a place with a high elevation.

Visualizing the distributions of temperatures and elevations
```{r}
barplot(table(met$temp))
barplot(table(met$elev))

hist(met$atm.press)
```

```{r}
plot(met$elev ~ met$atm.press)
```

I would have expected the highest pressure to be at the highest elevation, but that does not seem to be the case.